---
title: "Forecasting U.S. Gasoline Prices under Geopolitical Shocks"
author: "Team ShohanXuKimKim"
date: 04-10-2026
format:
  html:
    toc: true
    theme: cosmo
execute:
  warning: false
  message: false
editor: visual
---

## 1. Project Overview

We want to analyze how major geopolitical shocks affect U.S. gasoline prices and compare different time series models to see which performs best under these conditions.

-   **Russia–Ukraine War** (Feb 2022)
-   **Israel–Iran Conflict** (June 2025)
-   **US Intervention in Venezuela** (Jan 2026)

## 2. Data Loading & Preprocessing

To analyze the relationship between U.S. gasoline prices, crude oil prices, and gasoline stocks, we collected weekly data from the U.S. Energy Information Administration (EIA).

During the initial data loading phase, we identified a structural reporting discrepancy: weekly gasoline prices are recorded on **Mondays**, whereas WTI crude oil prices and gasoline stocks are reported on **Fridays**. To align the datasets correctly for our time series join, we advanced the WTI and stocks dates by 3 days.

After successfully merging the data and filtering the timeline to start from 2018, we applied a rigorous data cleaning process. We utilized linear interpolation to impute any missing values and applied STL decomposition-based smoothing to handle extreme outliers. This ensures that our forecasting models are robust and not skewed by temporary noise.

```{r}
# Load required packages
library(tidyverse)
library(lubridate)
library(timetk)

# Guys, Please type install.packages(c("tidymodels", "modeltime", "prophet", "forecast", "xgboost", "ranger", "skimr", "GGally")) in the console. (To be deleted)

# ==========================================
# 1. Read Data and Align Dates
# ==========================================
# Note: EIA CSV files contain 4 lines of metadata at the top, so we use skip = 4.

# Gas prices are reported on Mondays.
gas_price_df <- read_csv("../Raw_Data/Weekly U.S. Gas Price (EIA)", skip = 4) %>%
  rename(date = 1, gas_price = 2) %>% 
  mutate(date = mdy(date)) 

# WTI prices and Gas stocks are reported on Fridays. 
# We add 3 days to align them with the Monday schedule of Gas prices.
wti_price_df <- read_csv("../Raw_Data/Weekly WTI Price (EIA)", skip = 4) %>%
  rename(date = 1, wti_price = 2) %>% 
  mutate(date = mdy(date) + days(3)) 

gas_stocks_df <- read_csv("../Raw_Data/Weekly U.S. Gas Stocks (EIA)", skip = 4) %>%
  rename(date = 1, gas_stocks = 2) %>% 
  mutate(date = mdy(date) + days(3)) 

# ==========================================
# 2. Merge and Clean Data
# ==========================================
clean_data <- gas_price_df %>%
  left_join(wti_price_df, by = "date") %>%
  left_join(gas_stocks_df, by = "date") %>%
  arrange(date) %>% # Ensure chronological order
  filter(date >= as.Date("2018-01-01")) %>%
  mutate(
    # Step 1: Impute missing values (NAs) using interpolation
    # Step 2: Clean and smooth extreme anomalies out of the time series
    gas_price  = ts_clean_vec(ts_impute_vec(gas_price, period = 1), period = 1),
    wti_price  = ts_clean_vec(ts_impute_vec(wti_price, period = 1), period = 1),
    gas_stocks = ts_clean_vec(ts_impute_vec(gas_stocks, period = 1), period = 1)
  )

# Preview the cleaned and aligned data
head(clean_data)

# ==========================================
# 3. Export Processed Data
# ==========================================
# We save the cleaned and merged dataset for reproducibility and team sharing.
# The file is saved in CSV format within the 'Processed_Data' folder.

write_csv(clean_data, "../Processed_Data/cleaned_data.csv")
```

## 3. Exploratory Data Analysis (EDA)

Before fitting our forecasting models, it is essential to understand the underlying statistical properties of our dataset, including the individual trajectories of our variables, their correlation, and their structural components.

### Summary Statistics

We first inspect the summary statistics and the historical paths of our primary variables: Gasoline Price, WTI Crude Oil Price, and Gasoline Stocks.

```{r}
library(skimr)
library(GGally) # For correlation plots

# 1. Summary Statistics
skim(clean_data)

# 2. Plotting Each Variable
clean_data %>%
  pivot_longer(cols = c(gas_price, wti_price, gas_stocks), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = date, y = Value, color = Variable)) +
  geom_line() +
  facet_wrap(~ Variable, scales = "free_y", ncol = 1) +
  theme_minimal() +
  labs(title = "Historical Trajectories of Key Variables", x = "Date", y = "") +
  theme(legend.position = "none")
```

### Correlation Analysis

To justify the inclusion of WTI prices and gasoline stocks as external regressors in our ARIMAX and Prophet models, we evaluate their linear relationships.

```{r}
# Pearson Correlation Matrix Plot
clean_data %>%
  select(gas_price, wti_price, gas_stocks) %>%
  ggpairs(title = "Correlation Matrix: Gas Price vs External Regressors") +
  theme_bw()
```

### Visualization

And we visualize the cleaned gasoline prices.

```{r}
# Plot the target variable
clean_data %>%
  plot_time_series(date, gas_price, 
                   .interactive = TRUE, 
                   .title = "Cleaned U.S. Gasoline Prices")
```

## 4. Traditional ARIMA and SARIMA

### 4.1. Univariate ARIMA

we first analyze the univariate time series of gasoline prices. We will decompose the series, check for stationarity, manually test various $ARIMA(p,d,q)$ specifications, and select the optimal model based on the AIC criterion.

```{r}
# Load required traditional time series packages
library(forecast)
library(tseries)
library(cowplot)

# ==========================================
# 4.1 Transform to Time Series Object
# ==========================================
# Converting to a ts object. Assuming weekly data (frequency = 52).
ts_gas_price <- ts(clean_data$gas_price, start=c(2018, 1), frequency=52)

# ==========================================
# 4.2 Decompose and Remove Seasonality
# ==========================================

decompose_gas <- decompose(ts_gas_price, "additive")
plot(decompose_gas)

# Create non-seasonal series for ARIMA fitting
deseasonal_gas <- seasadj(decompose_gas)

# Comparing ACFs: Original vs Deseasonalized
plot_grid(
  autoplot(Acf(ts_gas_price, lag = 40, plot=FALSE), 
           main = "Original Gas Price"),
  autoplot(Acf(deseasonal_gas, lag = 40, plot=FALSE),  
           main = "Deseasonalized Gas Price")
)
```

### 4.2. Stationarity Test and Differencing

Before fitting, we use the Augmented Dickey-Fuller (ADF) test to verify stationarity.

#### original gas price

Since the ADF test p-value is larger than 0.05, we apply first differencing (d=1) and check ACF/PACF of differenced series. Since the graphy shows strong seasonality, we deseasonalize the data as a next step.

```{r}
# Run ADF Test
print(adf.test(gas, alternative="stationary"))

# If p-value > 0.05, it indicates a stochastic trend. We apply first differencing (d=1).
gas_diff <- diff(gas, differences=1)

# Check ACF/PACF of the differenced series to identify p and q

plot_grid(
  autoplot(Acf(gas_diff, lag = 40, plot=FALSE), 
           main = "ACF: Differenced Series (original)"),
  autoplot(Pacf(gas_diff, lag = 40, plot=FALSE),  
           main = "PACF: Differenced Series (original)")
)
```

#### deseasonalizaed gas price

After applying first differencing (d=1) to achieve stationarity, we examine the ACF and PACF plots to identify the p and q parameters.

The PACF plot displays a sharp cut-off after Lag 1, while the ACF plot shows a gradual decay (tailing off). In time series theory, this specific combination is the classic signature of an Autoregressive process of order 1, or AR(1). Therefore, the visual evidence strongly suggests an ARIMA(1,1,0) specification.

```{r}
# Run ADF Test
print(adf.test(deseasonal_gas, alternative="stationary"))

# If p-value > 0.05, it indicates a stochastic trend. We apply first differencing (d=1).
deseasonal_gas_diff <- diff(deseasonal_gas, differences=1)

# Check ACF/PACF of the differenced series to identify p and q

plot_grid(
  autoplot(Acf(deseasonal_gas_diff, lag = 40, plot=FALSE), 
           main = "ACF: Differenced Series (deseasonal)"),
  autoplot(Pacf(deseasonal_gas_diff, lag = 40, plot=FALSE),  
           main = "PACF: Differenced Series (deseasonal)")
)
```

### 4.3. Manual ARIMA Fitting & AIC Comparison

We manually test several combinations of Autoregressive ($p$) and Moving Average ($q$) terms. To select the most robust model, we evaluate both the Akaike Information Criterion (AIC) and the Bayesian Information Criterion (BIC), which applies a stricter penalty for model complexity. The result shows that ARIMA(1,1,0) is the strongest model.

```{r}
# Initialize an empty data frame to store AIC and BIC values
compare_criteria <- data.frame(Model=character(), AIC=numeric(), BIC=numeric(), stringsAsFactors=FALSE)

# Model 1: ARIMA(1,1,0)
Model_110 <- Arima(deseasonal_gas, order=c(1,1,0), include.drift=TRUE)
compare_criteria <- rbind(compare_criteria, data.frame(Model="ARIMA(1,1,0)", AIC=Model_110$aic, BIC=Model_110$bic))

# Model 2: ARIMA(1,1,1)
Model_111 <- Arima(deseasonal_gas, order=c(1,1,1), include.drift=TRUE)
compare_criteria <- rbind(compare_criteria, data.frame(Model="ARIMA(1,1,1)", AIC=Model_111$aic, BIC=Model_111$bic))

# Model 3: ARIMA(2,1,1)
Model_211 <- Arima(deseasonal_gas, order=c(2,1,1), include.drift=TRUE)
compare_criteria <- rbind(compare_criteria, data.frame(Model="ARIMA(2,1,1)", AIC=Model_211$aic, BIC=Model_211$bic))

# Model 4: ARIMA(0,1,1)
Model_011 <- Arima(deseasonal_gas, order=c(0,1,1), include.drift=TRUE)
compare_criteria <- rbind(compare_criteria, data.frame(Model="ARIMA(0,1,1)", AIC=Model_011$aic, BIC=Model_011$bic))

# Print manual comparisons sorted by AIC
print(compare_criteria %>% arrange(AIC))
```

### 4.4. Auto ARIMA and Residual Diagnostics

Finally, we use the auto.arima() algorithm to find the mathematically optimal specification and check its residuals for white noise using the Ljung-Box test. When we put deseasonal data into auto-fitting ARIMA model, the strongest model is ARIMA(1,1,0)(0,0,1)\[52\].

```{r}
# Auto-fitting ARIMA
Model_autofit <- auto.arima(deseasonal_gas, approximation = TRUE)
print(Model_autofit)

# Add to comparison
compare_criteria <- rbind(compare_criteria, data.frame(Model="Auto ARIMA", AIC=Model_autofit$aic, BIC=Model_autofit$bic))

# Print final criteria table
print(compare_criteria %>% arrange(AIC))

# Check residuals for the best model
checkresiduals(Model_autofit)
```

### 4.5. ARIMA vs SARIMA

We analyze the univariate time series of gasoline prices. We will contrast a standard ARIMA model on deseasonalized data with a Seasonal ARIMA (SARIMA) model on the original data.

#### Automated Differencing Tests (ndiffs & nsdiffs)

Instead of relying solely on visual inspection, we utilize statistical tests to determine the required number of non-seasonal (d) and seasonal (D) differences.

```{r}
# Find non-seasonal differencing (trend)
n_diff <- ndiffs(deseasonal_gas)
cat("Number of non-seasonal differencing needed (d): ", n_diff, "\n")

# Find seasonal differencing
ns_diff <- nsdiffs(ts_gas_price)
cat("Number of seasonal differencing needed (D): ", ns_diff, "\n")

# Perform differencing based on test results (assuming d=1, D=1 for weekly seasonality)
gas_trend_diff <- diff(ts_gas_price, lag=1, differences=1)
gas_both_diff  <- diff(gas_trend_diff, lag=52, differences=1)
```

#### ACF/PACF of Twice-Differenced Series

To identify the SARIMA orders (p,d,q)(P,D,Q)\[52\], we analyze the series that has been differenced for both trend and seasonality.

Analyzing the twice-differenced series allows us to visually identify the autoregressive (AR) and moving average (MA) components.

1.  Non-Seasonal Components (p,q): The PACF displays a sharp cut-off after the first lag, while the ACF exhibits a gradual, tailing-off decay over the initial lags. In time series theory, this pattern is the classic signature of an Autoregressive process. Thus, we identify p=1 and q=0.

2.  Seasonal Components (P,Q): Observing the seasonal lags (multiples of 52), the PACF shows a distinct spike at Lag 52 and cuts off, whereas the ACF decays more slowly across the seasonal periods (Lag 52, 104). This implies a seasonal Autoregressive process. Thus, we identify P=1 and Q=0.

Combining these observations with our known differencing orders (d=1,D=1), the visual evidence points us toward a SARIMA(1,1,0)(1,1,0)\[52\] specification.

```{r}
# Plot ACF and PACF for twice-differenced series
plot_grid(
  autoplot(Acf(gas_both_diff, lag.max=110, plot=FALSE), 
           main="ACF: Twice-Differenced", ylim=c(-1,1)),
  autoplot(Pacf(gas_both_diff, lag.max=110, plot=FALSE), 
           main="PACF: Twice-Differenced", ylim=c(-1,1)),
  nrow=1, ncol=2
)
```

#### Automatically Fitting SARIMA (Original) vs ARIMA (Deseasonalized)

We fit an automated SARIMA model to the original data to capture both dynamics simultaneously, and compare its residual performance against the non-seasonal ARIMA.

```{r}
# 1. SARIMA on Original Data
SARIMA_autofit <- auto.arima(ts_gas_price)
print(SARIMA_autofit)

# 2. ARIMA on Deseasonalized Data
ARIMA_autofit <- auto.arima(deseasonal_gas, max.D=0, max.P=0, max.Q=0)
print(ARIMA_autofit)

# Residual Comparison
plot_grid(
  autoplot(Acf(SARIMA_autofit$residuals, lag.max=60, plot=FALSE), main="ACF: SARIMA Residuals"),
  autoplot(Acf(ARIMA_autofit$residuals, lag.max=60, plot=FALSE), main="ACF: ARIMA Residuals"),
  nrow=1, ncol=2
)
```

### 4.6 Methodological Insight: The Paradox of Auto-ARIMA Selection

Interestingly, auto.arima() selected a non-seasonal ARIMA(1,1,0) for the raw data, while selecting a seasonal ARIMA(1,1,0)(0,0,1)[52] for the deseasonalized data. This apparent paradox is a well-known phenomenon in high-frequency time series:

In the raw data, the high variance and the massive AICc penalty associated with estimating 52-lag parameters outweighed the benefits of modeling seasonality, leading the algorithm to favor a simpler AR(1) process.

Conversely, the decompose() function leaves behind faint residual seasonal autocorrelations. Because the overall variance was drastically reduced in the deseasonalized series, the algorithm detected this subtle "seasonal ghost" and efficiently captured it using a single seasonal moving average term (Q=1).

This highlights the limitations of traditional fixed decomposition and confirms our rationale for transitioning to machine learning models (like Prophet) that can handle complex, evolving weekly seasonality without incurring excessive parameter penalties.

## 5. Machine-learning Methods

### 5.1. Feature Engineering & Train/Test Split

We will create dummy variables to represent our three major geopolitical shocks. Then, we will split the data, reserving the most recent 10% (approx. 26 weeks) for model testing.

```{r}
# Define shock dates
shock_ukraine <- as.Date("2022-02-24")
shock_israel_iran <- as.Date("2025-06-01") # June 2025
shock_venezuela <- as.Date("2026-01-01")   # Jan 2026

# Add dummy variables to the cleaned data
model_data <- clean_data %>%
  mutate(
    shock_ukraine_dummy = ifelse(date >= shock_ukraine, 1, 0),
    shock_israel_iran_dummy = ifelse(date >= shock_israel_iran, 1, 0),
    shock_venezuela_dummy = ifelse(date >= shock_venezuela, 1, 0)
  )

# Time-based split: 90% Training, 10% Testing
library(tidymodels)
library(modeltime)

splits <- initial_time_split(model_data, prop = 0.9)
```

### 5.2. Expanded Modeling Setup

To thoroughly evaluate the impact of geopolitical shocks, we expand our forecasting roster to include 7 distinct models. We structure our approach logically, progressing from a simple baseline to traditional statistical models, machine learning algorithms, and finally, complex neural networks.

```{r}
# 1. Seasonal Naive (Simplest Baseline)
model_snaive <- naive_reg() %>%
  set_engine("snaive") %>%
  fit(gas_price ~ date, data = training(splits))

# 2. ETS (Error, Trend, Seasonality)
model_ets <- exp_smoothing() %>%
  set_engine("ets") %>%
  fit(gas_price ~ date, data = training(splits))

# 3. ARIMAX (Traditional Statistical Model)
model_arima <- arima_reg() %>%
  set_engine("auto_arima") %>%
  fit(gas_price ~ date + wti_price + gas_stocks + shock_ukraine_dummy + shock_israel_iran_dummy + shock_venezuela_dummy, 
      data = training(splits))

# 4. Prophet (Generalized Additive Model)
model_prophet <- prophet_reg() %>%
  set_engine("prophet") %>%
  fit(gas_price ~ date + wti_price + gas_stocks + shock_ukraine_dummy + shock_israel_iran_dummy + shock_venezuela_dummy, 
      data = training(splits))

# 5. Random Forest (Machine Learning)
# Added .set_mode("regression") to specify we are predicting continuous values
model_rf <- rand_forest(trees = 500, min_n = 5) %>%
  set_engine("ranger") %>%
  set_mode("regression") %>% 
  fit(gas_price ~ date + wti_price + gas_stocks + shock_ukraine_dummy + shock_israel_iran_dummy + shock_venezuela_dummy, 
      data = training(splits))

# 6. ARIMA-Boost (Hybrid/Ensemble Model)
# Added .set_mode("regression") for the XGBoost component
model_arima_boost <- arima_boost() %>%
  set_engine("auto_arima_xgboost") %>%
  set_mode("regression") %>% 
  fit(gas_price ~ date + wti_price + gas_stocks + shock_ukraine_dummy + shock_israel_iran_dummy + shock_venezuela_dummy, 
      data = training(splits))

# 7. NNAR (Neural Network Auto-Regression)
model_nnar <- nnetar_reg() %>%
  set_engine("nnetar") %>%
  fit(gas_price ~ date + wti_price + gas_stocks + shock_ukraine_dummy + shock_israel_iran_dummy + shock_venezuela_dummy, 
      data = training(splits))
```

### 5.3. Comprehensive Model Evaluation

We consolidate all 7 models into a single table to benchmark their forecasting accuracy (e.g., RMSE, MAE, MAPE) on the withheld test dataset.

```{r}
# Create a Modeltime Table maintaining the simple-to-complex order
models_tbl <- modeltime_table(
  model_snaive,
  model_ets,
  model_arima,
  model_prophet,
  model_rf,
  model_arima_boost,
  model_nnar
)

# Calibrate models on the testing data
calibration_tbl <- models_tbl %>%
  modeltime_calibrate(new_data = testing(splits))

# Extract and display the accuracy metrics (Sorted by RMSE)
accuracy_tbl <- calibration_tbl %>%
  modeltime_accuracy() %>%
  arrange(rmse) 

print(accuracy_tbl)

# Plot Forecasts vs Actuals for all models
calibration_tbl %>%
  modeltime_forecast(
    new_data    = testing(splits),
    actual_data = clean_data
  ) %>%
  plot_modeltime_forecast(
    .legend_max_width = 25,
    .interactive      = TRUE,
    .title = "Forecast vs Actual: Model Comparison"
  )
```

Note that while manual specification with an annual frequency (52) omitted seasonal terms due to high AIC penalties, our later machine-learning-based modeltime approach intelligently captures quarterly seasonality (m=13). This perfectly illustrates why advanced, flexible frameworks are necessary for complex high-frequency data.

### 5.4. Visualization of Forecast Results (Full version)

Visualizing the forecast against the actual test data is crucial for understanding how each model reacts to the structural breaks caused by geopolitical shocks. The following plot allows us to compare the trajectory of our 7 models against the observed gasoline prices.

```{r}
# ==========================================
# 1. Generate and Plot the Forecast
# ==========================================
# We use the calibrated table to project our models onto the testing period.
# The .interactive = TRUE argument allows us to zoom in on specific shock periods.

calibration_tbl %>%
  modeltime_forecast(
    new_data    = testing(splits),
    actual_data = clean_data
  ) %>%
  plot_modeltime_forecast(
    .legend_max_width = 25, # Adjust legend width for 7 models
    .interactive      = TRUE,
    .conf_interval_show = FALSE, # Hide confidence intervals to reduce clutter
    .title = "Comparison of 7 Models: Forecasting U.S. Gasoline Prices",
    .x_lab = "Date",
    .y_lab = "Price (USD/Gallon)",
    .color_lab = "Model Type"
  )
```

```{r}
# ==========================================
# 2. Focused Visualization (Top Models)
# ==========================================
# Identify top models from the accuracy table and plot them individually 
# to inspect their fit during shock windows.

calibration_tbl %>%
  # Example: Selecting only the best-performing models to reduce visual noise
  filter(.model_id %in% c(4, 5, 6)) %>% 
  modeltime_forecast(
    new_data    = testing(splits),
    actual_data = clean_data
  ) %>%
  plot_modeltime_forecast(
    .interactive = TRUE,
    .title = "Deep Dive: Top Performing Models Under Geopolitical Stress"
  )
```

### 5.5. Visualization of Forecast Results (Focused version)

To capture the full context of the geopolitical shocks, starting with the Russia-Ukraine War in early 2022, we have focused this visualization on the period from 2022 to 2026. This allows us to see how each model responded to the initial major price spike and how they perform during the subsequent conflicts.

```{r}
# ==========================================
# 1. Plot Forecast - Focused from 2022
# ==========================================
# We use filter_by_time to zoom into the period starting from 2022.

calibration_tbl %>%
  modeltime_forecast(
    new_data    = testing(splits),
    actual_data = clean_data
  ) %>%
  # Filtering the plot to show data from January 2022 onwards
  filter_by_time(.date_var = .index, .start_date = "2022-01-01") %>%
  plot_modeltime_forecast(
    .legend_max_width = 25,
    .interactive      = TRUE,
    .conf_interval_show = FALSE,
    .title = "Comparison of 7 Models (From 2022)",
    .x_lab = "Date",
    .y_lab = "Price (USD/Gallon)"
  )
```

```{r}
# ==========================================
# 2. Focused Visualization (Top Models)
# ==========================================
# Selecting the best performing models (e.g., ARIMA-Boost, Prophet, RF)
# and restricting the timeline to 2022 onwards.

calibration_tbl %>%
  # Note: Adjust the .model_id indices based on your accuracy_tbl results
  filter(.model_id %in% c(4, 5, 6)) %>% 
  modeltime_forecast(
    new_data    = testing(splits),
    actual_data = clean_data
  ) %>%
  filter_by_time(.date_var = .index, .start_date = "2022-01-01") %>%
  plot_modeltime_forecast(
    .interactive = TRUE,
    .title = "Top Performing Models: Detailed View (From 2022)",
    .x_lab = "Date",
    .y_lab = "Price (USD/Gallon)"
  )
```

### 5.6. Refitting and Future Forecasting (Full version)

Now that we have evaluated the models, the final step is to refit them on the entire dataset (including the test set) to capture the most recent trends. We will then project the gasoline prices 12 weeks into the future.

For models that require external regressors (WTI prices, Stocks, and Shock Dummies), we assume that the crude oil prices and gasoline stocks will remain at their last observed levels for the duration of the forecast. The shock dummies will remain at 1, as we are still within the post-shock environment.

```{r}
# ==========================================
# 1. Refit Models to the Entire Dataset
# ==========================================
# Refitting ensures the models learn from the most recent 26 weeks of test data.
refit_tbl <- calibration_tbl %>%
  modeltime_refit(data = model_data)

# ==========================================
# 2. Create Future Data Frame (Next 12 Weeks)
# ==========================================
# We extend the timeline by 12 weeks and populate the external regressors.
future_data <- model_data %>%
  future_frame(.date_var = date, .length_out = "12 weeks") %>%
  # Assumption: Future external variables stay at the last known value
  mutate(
    wti_price = tail(model_data$wti_price, 1),
    gas_stocks = tail(model_data$gas_stocks, 1),
    shock_ukraine_dummy = 1,
    shock_israel_iran_dummy = 1,
    shock_venezuela_dummy = 1
  )

# ==========================================
# 3. Generate Future Forecast
# ==========================================
future_forecast_tbl <- refit_tbl %>%
  modeltime_forecast(
    new_data    = future_data,
    actual_data = model_data
  )

# ==========================================
# 4. Plot the Future Forecast
# ==========================================
future_forecast_tbl %>%
  plot_modeltime_forecast(
    .interactive = TRUE,
    .title = "12-Week Future Forecast: U.S. Gasoline Prices",
    .x_lab = "Date",
    .y_lab = "Price (USD/Gallon)"
  )
```

### 5.7. Refitting and 12-Week Future Forecast (Focused version)

In this final stage, we refit the models to the entire dataset, incorporating all available observations up to April 2026. We then project gasoline prices for the next 12 weeks. To clearly visualize the projected trends against the most recent market behavior, this plot focuses on the last 2 years (2024–2026).

```{r}
# ==========================================
# 1. Refit Models to the Full Dataset
# ==========================================
# This step ensures the models are updated with the latest trends from the test set.
refit_tbl <- calibration_tbl %>%
  modeltime_refit(data = model_data)

# ==========================================
# 2. Prepare Future Data Frame
# ==========================================
# Creating a 12-week horizon. External variables are assumed to hold 
# their last observed values.
future_data <- model_data %>%
  future_frame(.date_var = date, .length_out = "12 weeks") %>%
  mutate(
    wti_price = tail(model_data$wti_price, 1),
    gas_stocks = tail(model_data$gas_stocks, 1),
    shock_ukraine_dummy = 1,
    shock_israel_iran_dummy = 1,
    shock_venezuela_dummy = 1
  )

# ==========================================
# 3. Generate and Plot Future Forecast
# ==========================================
# We filter the visualization to start from 2024 to emphasize the 
# upcoming 12-week trajectory.
refit_tbl %>%
  modeltime_forecast(
    new_data    = future_data,
    actual_data = model_data
  ) %>%
  # Filtering the view to the last 2 years for better clarity
  filter_by_time(.date_var = .index, .start_date = "2024-01-01") %>%
  plot_modeltime_forecast(
    .interactive = TRUE,
    .title = "12-Week Future Forecast (Recent 2-Year Perspective)",
    .x_lab = "Date",
    .y_lab = "Price (USD/Gallon)"
  )
```

## 6. Event-Specific Performance Deep Dive (Event Study)

To rigorously test the robustness of our models, we conduct a "retrospective evaluation" for each of the three major geopolitical shocks. Instead of a global evaluation, we isolate the specific periods following each event to identify which models are most resilient to different types of structural breaks.

### 6.1. Shock 1: Russia-Ukraine War (Feb 2022)

We evaluate the forecast accuracy for the 12-week period immediately following the invasion.

```{r}
# 1. Define the Post-Shock Window for Ukraine
ukraine_shock_date <- as.Date("2022-02-24")

# 2. Filter data for a local evaluation (mimicking real-time testing)
# We use the already calibrated models but focus on the residuals during this specific window
ukraine_accuracy <- calibration_tbl %>%
  # Note: This requires the calibration to have covered this period. 
  # If the global test set is only at the end, we analyze the 'fitted' values here.
  modeltime_accuracy(
    new_data = model_data %>% filter(date >= ukraine_shock_date & date <= ukraine_shock_date + weeks(12))
  ) %>%
  arrange(rmse)

print(ukraine_accuracy)
```

### 6.2. Shock 2: Israel-Iran Conflict (June 2025)

This analysis focuses on the 12-week volatility period in mid-2025.

```{r}
# 1. Define the Post-Shock Window for Israel-Iran
israel_shock_date <- as.Date("2025-06-01")

# 2. Local Accuracy Evaluation
israel_accuracy <- calibration_tbl %>%
  modeltime_accuracy(
    new_data = model_data %>% filter(date >= israel_shock_date & date <= israel_shock_date + weeks(12))
  ) %>%
  arrange(rmse)

print(israel_accuracy)
```

### 6.3. Shock 3: US Intervention in Venezuela (Jan 2026)

As the most recent shock, this evaluation tests how models handle the latest structural shifts.

```{r}
# 1. Define the Post-Shock Window for Venezuela
venezuela_shock_date <- as.Date("2026-01-01")

# 2. Local Accuracy Evaluation
venezuela_accuracy <- calibration_tbl %>%
  modeltime_accuracy(
    new_data = model_data %>% filter(date >= venezuela_shock_date & date <= venezuela_shock_date + weeks(12))
  ) %>%
  arrange(rmse)

print(venezuela_accuracy)
```

## 7. Analysis of Model Robustness Across Geopolitical Shocks

Our event study reveals a critical finding: **there is no single "best" model for all types of geopolitical stress.** The predictive power of each algorithm shifted significantly depending on the nature of the shock.

1.  **Shock 1 (Russia-Ukraine): The Power of Ensembles** The **ARIMA-Boost** model outperformed all others with a remarkably low MAE (0.0143). This suggests that for massive, sudden price spikes, combining traditional statistical methods (ARIMA) with machine learning (XGBoost) is the most effective way to capture both the trend and the sudden residual volatility.

2.  **Shock 2 (Israel-Iran): Neural Network Superiority** During the mid-2025 volatility, the **NNAR (Neural Network Auto-Regression)** model took the lead (MAE: 0.0201). This indicates that the price fluctuations during this period had complex, non-linear patterns that standard linear models struggled to identify, but which the neural network was able to learn through its hidden layers.

3.  **Shock 3 (Venezuela Intervention): Robustness to Structural Breaks** The most recent shock in early 2026 proved to be the most challenging for all models, as evidenced by the higher MAE scores across the board. However, **Prophet with Regressors** emerged as the winner (MAE: 0.1107). Prophet’s design—which is specifically built to handle trend changepoints and outliers—made it the most resilient model during this latest period of uncertainty.

## 8. Conclusion for Policy Makers

Based on these findings, we recommend a **"Dynamic Ensemble Approach."** ?!
